Secure Composition \\ of Robust and Optimising Compilers

### 2024-05-23 csf2/main.tex
add some rules and text

\vspace{-1.5em}
\[
  \begin{array}{rcl}
    \obj{D} &:=& \obj{ON} \mid \obj{OFF} \hspace{2em}
    \obj{\securitytag} := \obj{\lock} \mid \obj{\unlock} \\
    \obj{\event[^b]} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \mid \obj{Binop\ n} \mid \obj{Branch\ n} \\
    \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \mid \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
    \obj{\event} &:=& \obj{\event[^b];\sandboxtag;\securitytag} \mid \dots
  \end{array}
\]

@@

@@

\Cref{thm:rtpsim:sig} is always applicable, up to well-formedness of $\sim_1$ with respect to $\trg{\class[_2]}$.
However, consider $\ccb$ and $\ccspec$.

>>>>

\vspace{-1.5em}
\[
  \begin{array}{rcl}
    \obj{D} &:=& \obj{ON} \mid \obj{OFF} \hspace{2em}
    \obj{\securitytag} := \obj{\lock} \mid \obj{\unlock} \\
    \obj{\event[^b]} &:=& \dots \mid \obj{iGet\ \loc\ v} \mid \obj{iSet\ \loc\ v} \mid \obj{Binop\ n} \mid \obj{Branch\ n} \\
    \obj{\expr} &:=& \dots \mid \obj{\lrddoit{x}{\expr}} \mid \obj{\lwrdoit{D}} \mid \obj{\llet{x^{\securitytag}}{\expr[_1]}{\expr[_2]}}\\
    \obj{\event} &:=& \obj{\event[^b];\sandboxtag;\securitytag} \mid \dots
  \end{array}
\]
\begin{center}
\footnotesize
  \typerule{$e-\obj{wrdoit}-\text{off}$}{
  }{
    \pstepto{\rtt{\obj{\cfstate;\sandboxtag;m;\memstate}}{\obj{\lwrdoit{OFF}}}}{\rtt{\obj{\cfstate;\sandboxtag;OFF;\memstate}}{\obj{0^{\securitytag}}}}{\obj{\emptyevent}}
  }{e-wrdoit-off}
\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
\newcommand{\exprevalo}[5]{\expreval{\obj{#1}}{\obj{#2}}{\obj{#3}}{\obj{#4}}{\obj{#5}}}
  \typerule{$e-\obj{get}-\in-$noleak}{
    \obj{\memstate}=\obj{H^{\ctx};H^{\comp};\Delta_1,x\mapsto(\loc;\sandboxtag;\poison;m),\Delta_2} &
    \obj{\loc}+\obj{n}\in\text{dom }\obj{H^{\sandboxtag}} \\
    \obj{\sigma''}=\obj{\sigma\sqcap\sigma'}
  }{
    \exprevalo{\cfstate;\sandboxtag';m+1;\memstate}{x^\sigma[n^{\sigma'}]}{\cfstate;\sandboxtag';m+1;\memstate}{(H^{\sandboxtag}(\loc+n))^{\sigma''}}{{iGet}\ \loc\ n;\sandboxtag;\sigma''}
  }{to-e-get-in-noleak}
\end{center}

@@

\begin{center}
  \footnotesize
\newcommand{\expreval}[5]{{#1}\triangleright\xspace {#2}\xrightarrow{#5}\ {#3}\triangleright\xspace {#4}\xspace}
\newcommand{\exprevald}[5]{\expreval{\ird{#1}}{\ird{#2}}{\ird{#3}}{\ird{#4}}{\ird{#5}}}

  %
  \typerule{$e-\ird{ifz}-\ird{true}-\text{spec}$}{
    \dots \text{todo}
  }{
    \pstepto{\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\lifz{0^{\securitytag}}{\expr[_1]}{\expr[_2]}}}}{\begin{array}{c}\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_1]}}\\\rtt{\ird{\cfstate;\sandboxtag;0;\memstate}}{\ird{\expr[_2]}}\end{array}}{\ird{\ev{Branch}\ 0;\sandboxtag;\securitytag}\cdot\ird{\ev{Spec};\sandboxtag;\securitytag}}
  }{e-ifz-true-leak}
\end{center}

@@

\Cref{thm:rtpsim:sig} is always applicable, up to well-formedness of $\sim_1$ with respect to $\trg{\class[_2]}$.
However, consider $\ccb$ and $\ccspec$.
Even though swapping these compilers in the compilation chain is syntactically impossible, let us pretend that it is possible by just extending $\ccb$ for the constructs present in $\Lscct$ that are not present in $\Ltrg$.
Clearly, as evidenced in the case study earlier, $\ccb\cdot\ccspec$ is an acceptable composition that yields an interesting class of security properties, i.e., $\sigma_{\sim_{\irl{ms}}\bullet\sim_{\ird{\mathghost}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$. 
However, swapping them $\ccspec\cdot\ccb$ means that $\ccb$ inserts new branches into the code that are not necessarily protected by a speculative execution barrier! 
It is crucial to ensure that the resulting class is meaningful, much similar to ensuring that a definition models what the formal methods expert intends to model. 
For the $\ccspec\cdot\ccb$ case, the class is $\sigma_{\sim_{\ird{\mathghost}}\bullet\sim_{\irl{ms}}}\left(\mssafe\cap\scctsafe\cap\sssafe\right)$
Since $\ev{Use\ \loc\ n}\sim_{\irl{ms}}\ev{Branch n}\cdot\ev{Spec}\cdots$, where $\cdots$ does not contain a $\ird{\ev{Barrier}}$ event, the resulting class of properties is only a subset of the intended one.
Because of this, it is crucial to analyze the precise shape of the class of properties after mapping a target-level property to a source-level property\footnote{Dito for vice-versa.}.
